---
title: "The Flexible Gumbel Distribution"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is the GitHub repository for the paper titled "The Flexible Gumbel Distribution: A New Model for Inference About the Mode" (**add the link later**). In this paper, we introduce a new unimodal distribution indexed by the mode. This distribution is a mixture of Gumbel distributions for the maximum and minimum values, which we refer to as the flexible Gumbel distribution. 

The probability density function (pdf) of the flexible Gumbel distribution is given by:

$$
\begin{aligned}
f(y)= & w \times \frac{1}{\sigma_1} \exp [-\frac{x-\theta}{\sigma_1}-\exp (-\frac{x-\theta}{\sigma_1})]+ \\
& (1-w) \times \frac{1}{\sigma_2} \exp [\frac{x-\theta}{\sigma_2}-\exp (\frac{x-\theta}{\sigma_2})],
\end{aligned}
$$

where $w \in[0,1]$ represents the mixing proportion parameter. We denote that $Y$ follows the distribution specified by the pdf as $Y \sim \text{FG}(\theta, \sigma_1, \sigma_2, w)$.

## Sampling from the Flexible Gumbel Distribution

Sampling from the flexible Gumbel distribution is straightforward due to its mixture distribution nature with two components. To generate $n$ independent and identically distributed (i.i.d.) samples from the flexible Gumbel distribution, you can utilize the provided functions in this GitHub repository. Below is an example of sampling 1000 i.i.d. samples from $\text{FG}(0,1,1,0.3)$.

```{r, fig.width=7, fig.height=3.5, fig.align='center'}
source("./density.R")

set.seed(100)

# Set the sample size
n <- 1000

# Define parameters
w1 <- 0.2
loc <- 0
scale1 <- 1
scale2 <- 1

# Sample from the mixture distribution
Z <- rbinom(n=n, size=1, prob=w1)
X1 <- rgumbel_RNG(n=n, loc, scale1)
X2 <- 2 * loc - rgumbel_RNG(n=n, loc, scale2)
Y <- Z * X1 + (1 - Z) * X2

# Plot histogram and density
hist(Y, breaks=50, main="Distribution of FG Samples", freq=FALSE)
points(density(Y), type="l", col=rgb(0.8, 0, 0, 1))

# Add density curve of the true pdf
xaxis <- seq(min(Y)*1.2,max(Y)*1.2,length.out = 1000)
yaxis <- mixture_gumbel_pdf(x = xaxis, w1 = w1, loc = loc, 
                            scale1 = scale1, scale2 = scale2, log = FALSE)
lines(xaxis, yaxis, type="l", col=rgb(0, 0.8, 0, 1), lty=2)

# Add legend
legend("topright", legend=c("Sample Density", "True Density"), 
       col=c("red", "green"), lty=c(1, 2), cex=0.8)
```

## Frequentist Inference

The Frequentist inference of the flexible Gumbel (FG) distribution includes an implementation of maximum likelihood estimation (MLE) and the associated sandwich variance estimation to quantify the uncertainty of the MLE. We utilized the Expectation–Conditional Maximization (ECM) algorithm to calculate the MLE estimation of the FG distribution. An example of how to use the ECM algorithm program is provided below.

```{r, message=FALSE, warning=FALSE}
source("./EM_v1.0.R")
estimation <- em(Y, w1, loc, scale1, scale2)
print(round(estimation,3))
SE <- sandwichvar_se(Y, estimation$w1, estimation$loc, estimation$scale1, estimation$scale2)
print(round(SE,3))
```

## Bayesian Inference

This project includes an implementation of Markov Chain Monte Carlo (MCMC) for Bayesian inference of the FG distribution. We employ the Metropolis-Hastings within Gibbs algorithm to draw samples from the posterior distribution. Below is an example demonstrating how to conduct Bayesian inference for the FG distribution using our provided functions.

```{r, message=FALSE, warning=FALSE}
source("./Bayes_MH_v4.0.r")

# Set initial parameter values
w1_initial <- 0.3
loc_initial <- 0
scale1_initial <- 1
scale2_initial <- 1

# Tune the proposal distribution parameters
loc_tune <- 0.1
scale1_tune <- 0.2
scale2_tune <- 0.1

# Run MCMC sampling
output <- MH(data = Y,
             w1_initial = w1_initial,
             loc_initial = loc_initial,
             scale1_initial = scale1_initial,
             scale2_initial = scale2_initial,
             loc_tune = loc_tune,
             scale1_tune = scale1_tune,
             scale2_tune = scale2_tune,
             random_seed = 100,
             burn_in = 5000,
             iteration = 10000,
             thin = 1)

estimation <- data.frame(t(round(summary(output)[[2]][, 3], 3)))
print(estimation)
SE <- round(summary(output)[[1]][, 2], 3)
print(SE)
```

## R shiny app for the Flexible Gumbel Distribution

We have deployed an [R Shiny app](https://qingyang.shinyapps.io/gumbel_mixture/) for the FG distribution. This interactive tool allows users to explore various aspects of the FG distribution, including:

- Generating PDF plots and cumulative density plots for the FG distribution with different location/mode, scale, and weight parameters.
- Performing calculations and visualizations to verify the identifiability of the FG distribution.
- Calculating third and fourth central moments (skewness and kurtosis) of the FG distribution.
- Conducting both Frequentist and Bayesian inferences using uploaded CSV files.

The [R Shiny app](https://qingyang.shinyapps.io/gumbel_mixture/) provides a user-friendly interface for conducting analyses and visualizations related to the FG distribution. Feel free to explore and utilize the app for your statistical needs.

## `STAN` and `JAGS` Programs

The repository includes `STAN` and `JAGS` programs for implementing the Flexible Gumbel (FG) distribution:

- The `STAN` program can be accessed [here](https://github.com/rh8liuqy/flexible_Gumbel/tree/main/table2/model_fitting.stan).
- The `JAGS` program can be accessed [here](https://github.com/rh8liuqy/flexible_Gumbel/tree/main/table2/Gumbel_Mixture_JAGS.BUG).

These programs are useful for conducting Bayesian inference and fitting the FG distribution to data using the `STAN` and `JAGS` software platforms, respectively. Feel free to utilize these programs for your statistical analyses.
